---
title: "General envPred usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{General envPred usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message = FALSE}
library(envPred)
```

First, we recommend carefully reading the help page for the main functions

```{r, eval = FALSE}
?predictability
?colwell74
```

To calculate seasonality and colour of environmental noise using sample data (temporally even SST data):

```{r}
sst_pred <- predictability(sst$time_series, sst$dates, delta = 1,
                           is_uneven = FALSE, interpolate = FALSE,
                           show_warns = TRUE, noise_method = "spectrum")
```

The package threw a warning, which can be addressed in this example by:

```{r, warnings = FALSE}
sst  <-  sst[sst$dates <= as.Date("2007-01-01"), ]
sst_pred <- predictability(sst$time_series, sst$dates, delta = 1,
                           is_uneven = FALSE, interpolate = FALSE,
                           show_warns = TRUE, noise_method = "spectrum")
```

The data can be plotted using [`ggplot2`](https://ggplot2.tidyverse.org/) methods. The use can plot both the de-trended dataset,

```{r}
gg_envpred(sst_pred, type = "detrended")
```

or the spectral density on the log 10 scale

```{r}
gg_envpred(sst_pred, type = "spectral")
```

Finally, the package also handles temporally uneven NPP data
with NAs (carefully check warnings), by using the Lomb-Scargle Periodogram from package
[`lomb`](https://CRAN.R-project.org/package=lomb).

```{r, warning = FALSE}
npp_pred  <-  predictability(npp$time_series, npp$dates, delta = 8,
                             is_uneven = TRUE, interpolate = TRUE,
                             show_warns = TRUE, noise_method = "lomb_scargle")
```

To obtain Colwell's metrics, simply run:

```{r}
colwell74(sst$time_series, sst$dates, n_states = 11)
```

The argument `n_states` is a numeric vector of length 1 containing a somewhat arbitrary number, as Colwell's method divides a continuous variable up into discrete states (read the original [paper](http://onlinelibrary.wiley.com/doi/10.2307/1940366/) for further details). Default (arbitrary) is 11.

The package also has a wrapper function that binds the output of `predictability` and `colwell74` together:

```{r, warning = FALSE}
all_pred_sst  <-  predictability_and_colwell(sst$time_series, sst$dates, n_states = 11, delta = 1,
                                             is_uneven = FALSE, interpolate = FALSE,
                                             show_warns = TRUE, noise_method = "spectrum")
```
